Well-quasi-orderings on word languages

### 2024-08-05 src/infixes.tex
adding monoid questions...
Adding automata based approach + theorem

In this section, we study the well-quasi-ordering of languages under the infix
relation.

\subsection{Lower bound}

Approximative theorem.
\begin{theorem}
    Let $(X, \preceq)$ be a quasi-ordered set.
    Then the following are equivalent
    \begin{itemize}
        \item $X$ embeds into $(\Sigma^*, \preceq_{\text{infix}})$
        \item $X$ is countable, and for every $x \in X$,
            $\{y \in X \mid y \preceq x\}$ is finite.
    \end{itemize}

@@

\subsection{Upper bound}

@@

>>>>

In this section, we study the well-quasi-ordering of languages under the infix
relation. As opposed to the prefix and suffix relations, the infix relation can
yield to very complicated well-quasi-ordered languages. As a matter of fact,
the following theorem shows that \emph{any} countable quasi-ordering with
finite initial segments can be embedded into the infix relation of a language.

\begin{theorem}
    \label{thm:infix-embedding}
    Let $(X, \preceq)$ be a quasi-ordered set.
    Then the following are equivalent
    \begin{itemize}
        \item $X$ embeds into $(\Sigma^*, \preceq_{\text{infix}})$
        \item $X$ is countable, and for every $x \in X$,
            $\{y \in X \mid y \preceq x\}$ is finite.
    \end{itemize}

@@

As a consequence of \cref{thm:infix-embedding}, we cannot replay proofs of
\cref{prefixes-suffixes:sec}, and will actually need to leverage some
regularity of the languages to obtain a characterization of well-quasi-ordered
languages under the infix relation. Let us first play this game for languages
that are recognized by finite automata.

\begin{theorem}
    \label{thm:infix-finite-automata}
    Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
    a finite union of chains for the \kl{infix relation}.
\end{theorem}

\begin{corollary}
    Decidability?
\end{corollary}

\subsection{Amalgamation systems}

@@

\subsection{Monoid equations?}

\begin{itemize}
    \item What are the finite monoids $(M, \cdot)$ such that
    every language recognized by $M$ is well-quasi-ordered by the infix relation?
    \item What are the equations on $M$ and $P \subseteq M$ characterizing 
    the well-quasi-ordering of $L(M, P)$?
\end{itemize}
